### 2017-07-13 draft/unique-linear.tex
finished adding joins and meets to typing
comments

  \inferrule
  {
    A = A_1 \OVERRIDE (A_2 \sqcup A_3) \\
    A_1 \vdash e : \NUM \\
    A_2 \vdash f : t_2 \\
    A_3 \vdash g : t_3 \\
    \SPLIT{t_1}{  t_2}{ t_3}
  }
  {
    A \vdash \IF e f g : t_1
  }
  % PT: is subtyping really needed? We are propagating demand information *backwards*, so (e.g.) the demand
  % on the branches of a conditional is the same as the demand on the result of the conditional.
  % \inferrule[Sub]
  % {A \vdash e : s \\ s \SUB t}
  % {A \vdash e : t}
  \end{mathpar}
  \caption{Typing rules.
  It seems like the subtyping in the FIX rule can alway be eliminated by unfolding the fixed point
  once.
  If the demand on $t$ is stronger or weaker than $s$, then unfolding will mediate the demand.
  NB the directionality of subtyping is wrong if interpolated from the lattice ordering (as it is presently).
}
  \label{fig:field-effects}
\end{figure}
% \begin{figure}[tp]
%   \begin{mathpar}
%     \inferrule{}{\NUM \SUB \NUM}
%
%     \inferrule{
%       s' \SUB s \\ t \SUB t'
%     }{
%       s \to t \SUB s' \to t' 
%     }
%
%     \inferrule{
%       \overline{s_i \SUB t_i} \\
%       \overline{P_i \sqsubseteq Q_i}
%     }{
%       \REC{ \overline{a_i :^{P_i} s_i}}
%       \SUB
%       \REC{ \overline{a_i :^{Q_i} t_i}}
%     }
%   \end{mathpar}
%   \caption{Subtyping}
%   \label{fig:subtyping-3}
% \end{figure}
\begin{figure}[tp]
  \begin{align*}
    \NUM \sqcup \NUM
    & = \NUM
    \\
    (s \to t) \sqcup (s' \to t')
    & = (s \sqcap s') \to (t \sqcup t') \qquad ???
    \\
    \REC{ \overline{a_i :^{P_i} s_i}} \sqcup \REC{ \overline{a_i :^{Q_i} t_i}}
    &= \REC{ \overline{a_i :^{P_i \sqcup Q_i} (s_i \sqcup t_i)}}
  \end{align*}
  \begin{mathpar}
    \inferrule{}{
      \SPLIT \NUM \NUM \NUM
    }

@@

    \inferrule{
      ??? \\
      \JOIN{s}{ s_1}{ s_2} \\
      \SPLIT{t}{t_1}{t_2}
    }{
      \JOIN{s \to t}{s_1 \to t_1}{s_2 \to t_2}
    }
    \\
    \inferrule
    {\overline{\SPLIT{s_i}{s_i'}{s_i''}}}
    {\SPLIT{\REC{ \overline{a_i :^{P_i} s_i}}}{\REC{ \overline{a_i :^{P_i} s_i'}}}{\REC{ \overline{a_i :^{P_i} s_i''}}}}
  \end{mathpar}
  \caption{Bounding and splitting operations on types}
  \label{fig:bounding-3}
\end{figure}
The novel ingredient in the typing rules are the propagation and splitting rules for the effect annotations on
record labels indicated by the ternary override relation $\OVERRIDE$.
Function closures have ``infinite'' splitting $(a :^P t)^*$ into $(a :^Q t)$.
\begin{itemize}
\item $P = \NOTHING$ propagate $Q=\NOTHING$.
\item $P = \READ$ propagate $Q = \WRITE$, no action required inside closure.
\item $P = \WRITE$ propagate $Q = \WRITE$, perform copy inside of closure.
\end{itemize}
 Propagating $\READ$ is not safe in both cases because that enables a write operation on the same
 record after the closure has been constructed. Alternatively, we could have another READONLY mode,
 which prevents subsequent write operations.

>>>>

  \inferrule
  {
    A = A_0 \OVERRIDE A_1 \\
    \JOIN{A_1}{A_2}{ A_3} \\
    \SPLIT{t_1}{  t_2}{ t_3} \\\\
    A_0 \vdash e : \NUM \\
    A_2 \vdash f : t_2 \\
    A_3 \vdash g : t_3
  }
  {
    A \vdash \IF e f g : t_1
  }
  % PT: is subtyping really needed? We are propagating demand information *backwards*, so (e.g.) the demand
  % on the branches of a conditional is the same as the demand on the result of the conditional.
  % \inferrule[Sub]
  % {A \vdash e : s \\ s \SUB t}
  % {A \vdash e : t}
  \end{mathpar}
  \caption{Typing rules.
  It seems like the subtyping in the FIX rule can alway be eliminated by unfolding the fixed point
  once.
  If the demand on $t$ is stronger or weaker than $s$, then unfolding will mediate the demand.
  NB the directionality of subtyping is wrong if interpolated from the lattice ordering (as it is presently).
}
  \label{fig:field-effects}
\end{figure}
% \begin{figure}[tp]
%   \begin{mathpar}
%     \inferrule{}{\NUM \SUB \NUM}
%
%     \inferrule{
%       s' \SUB s \\ t \SUB t'
%     }{
%       s \to t \SUB s' \to t' 
%     }
%
%     \inferrule{
%       \overline{s_i \SUB t_i} \\
%       \overline{P_i \sqsubseteq Q_i}
%     }{
%       \REC{ \overline{a_i :^{P_i} s_i}}
%       \SUB
%       \REC{ \overline{a_i :^{Q_i} t_i}}
%     }
%   \end{mathpar}
%   \caption{Subtyping}
%   \label{fig:subtyping-3}
% \end{figure}
\begin{figure}[tp]
  % \begin{align*}
  %   \NUM \sqcup \NUM
  %   & = \NUM
  %   \\
  %   (s \to t) \sqcup (s' \to t')
  %   & = (s \sqcap s') \to (t \sqcup t') \qquad ???
  %   \\
  %   \REC{ \overline{a_i :^{P_i} s_i}} \sqcup \REC{ \overline{a_i :^{Q_i} t_i}}
  %   &= \REC{ \overline{a_i :^{P_i \sqcup Q_i} (s_i \sqcup t_i)}}
  % \end{align*}
  \begin{mathpar}
    \inferrule{}{
      \SPLIT \NUM \NUM \NUM
    }

@@

    \inferrule{
      \SPLIT{s}{ s_1}{ s_2} \\
      \JOIN{t}{t_1}{t_2}
    }{
      \JOIN{s \to t}{s_1 \to t_1}{s_2 \to t_2}
    }
    \\
    \inferrule
    {\overline{\SPLIT{s_i}{s_i'}{s_i''}}}
    {\SPLIT{\REC{ \overline{a_i :^{P_i} s_i}}}{\REC{ \overline{a_i :^{P_i} s_i'}}}{\REC{ \overline{a_i :^{P_i} s_i''}}}}

    \inferrule
    {\overline{\JOIN{s_i}{s_i'}{s_i''}} \\
    \overline{O_I = P_i \sqcup Q_i}}
    {\JOIN{\REC{ \overline{a_i :^{O_i} s_i}}}{\REC{ \overline{a_i :^{P_i} s_i'}}}{\REC{ \overline{a_i :^{Q_i} s_i''}}}}
  \end{mathpar}
  \caption{Bounding and splitting operations on types. The intuition is that \SPLITK{} propagates
    information in the positive positions of a type downwards towards the subexpressions. At the
    same time, it joins information in the negative positions and propagates the join upwards. The
    \JOINK{} operation works dually.}
  \label{fig:bounding-3}
\end{figure}
The novel ingredient in the typing rules are the propagation and splitting rules for the effect annotations on
record labels indicated by the ternary override relation $\OVERRIDE$.
Function closures have ``infinite'' splitting $(a :^P t)^*$ into $(a :^Q t)$.
\begin{itemize}
\item $P = \NOTHING$ propagate $Q=\NOTHING$.
\item $P = \READ$ propagate $Q = \WRITE$, no action required inside closure.
\item $P = \WRITE$ propagate $Q = \WRITE$, perform copy inside of closure.
\end{itemize}
 Propagating $\READ$ is not safe in both cases because that enables a write operation on the same
 record after the closure has been constructed. Alternatively, we could have another READONLY mode,
 which prevents subsequent write operations.
